---
title: "p8105_mtp_sl4662"
output: github_document
editor_options: 
  chunk_output_type: console
---
# Problem 1

Some age group does not match the age. Age group 1 is not suppose to appear, and there is an age 45, with age group 1. 

```{r setup, include=FALSE}
devtools::install_github("thomasp85/patchwork")
library(tidyverse)
library(ggplot2)
library(rvest)
library(httr)
library(readxl)
library(patchwork)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "100%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
mtp_data = 
  read_excel("p8105_mtp_data.xlsx", 
             skip = 8) %>% 
  janitor::clean_names() %>% 
  mutate(
    eop_size_mm = replace(eop_size_mm, is.na(eop_size_mm), "0"), 
    eop_size_mm = as.double(eop_size_mm), 
    sex = factor(sex, levels = c("1", "0")), 
    sex = fct_recode(sex, female = "0", male = "1"), 
    age_group = factor(age_group, levels = c("1", "2", "3", "4", "5", "6", "7", "8")),  
    age_group = fct_recode(age_group, 
                           "18-30" = "2", 
                           "31-40" = "3", 
                           "41-50" = "4", 
                           "51-60" = "5", 
                           "60+" = "6", 
                           "60+" = "7", 
                           "60+" = "8"),
    eop_size = factor(eop_size, levels = c("0", "1", "2", "3", "4", "5")), 
    eop_size = fct_recode(eop_size, 
                          "0-5" = "0", 
                          "5-10" = "1", 
                          "10-15" = "2", 
                          "15-20" = "3", 
                          "20-25" = "4", 
                          "25+" = "5"), 
    eop_visibility_classification = factor(eop_visibility_classification, levels = c("0", "1", "2")), 
    eop_visibility_classification = fct_recode(eop_visibility_classification, 
                                               "eop size 0 mm" = "0",
                                               "0 < eop size mm <=5" = "1", 
                                               "eop size >= 5 mm" = "2"),
    fhp_category = factor(fhp_category, levels = c("0", "1", "2", "3", "4", "5", "6", "7")),
    fhp_category = fct_recode(fhp_category, 
                              "0-10" = "0", 
                              "10-20" = "1", 
                              "20-30" = "2", 
                              "30-40" = "3", 
                              "40-50" = "4", 
                              "50-60" = "5", 
                              "60-70" = "6", 
                              "70-80" = "7")
    )

  
# levels(mtp_data$age_group) = c("1", "20+", "30+", "40+", "50+", "60+", "60+", "60+")

age_female = 
  mtp_data %>% 
  filter(sex == "female") %>% 
  group_by(age_group) %>% 
  summarize(female_number = n())

age_male = 
  mtp_data %>% 
  filter(sex == "male") %>% 
  group_by(age_group) %>% 
  summarize(male_number = n())

left_join(age_female, age_male, by = "age_group") %>% 
  knitr::kable()
```

This dataset is consist of many different variables, and some variables have conclusive variables following them to classify the results into different groups. 

Key variables include :

There are `r ncol(mtp_data)` variables and `r nrow(mtp_data)` observations in this data set. There is a total of `r nrow(mtp_data)` participants. 


age = 
  mtp_data %>% 
  mutate(
    age = case_when(
      age < 18  ~ "<18",
      age > 17 & age < 31  ~ "18-30",
      age > 30 & age < 41  ~ "31-40",
      age > 40 & age < 51  ~ "41-50",
      age > 50 & age < 61  ~ "51-60",
      age > 60 ~ ">60",
      TRUE     ~ ""
  )) %>% 
  janitor::tabyl(sex, age) %>% 
  select("sex", "<18", "18-30","31-40","41-50","51-60",">60")

age_cat = 
  mtp_data %>% 
  mutate(
    age_cat = case_when(
      age_group == "1"  ~ "<18",
      age_group == "18-30"  ~ "18-30",
      age_group == "31-40"  ~ "31-40",
      age_group == "41-50"  ~ "41-50",
      age_group == "51-60"  ~ "51-60",
      age_group == "60+" ~ ">60",
    TRUE     ~ ""
  )) %>% 
  janitor::tabyl(sex, age_cat) %>% 
  select("sex", "<18", "18-30","31-40","41-50","51-60",">60")

aa = 
  bind_rows(age_cat, age) %>% 
  pivot_longer(
    names_to = female, 
              values_from = )

eop = 
  mtp_data %>% 
  mutate(
    eop = case_when(
      "eop_size mm" < 18  ~ "<18",
      age > 17 & age < 31  ~ "18-30",
      age > 30 & age < 41  ~ "31-40",
      age > 40 & age < 51  ~ "41-50",
      age > 50 & age < 61  ~ "51-60",
      age > 60 ~ ">60",
    TRUE     ~ ""
  )) %>% 
  janitor::tabyl(sex, age_cat) 



```{r, fig.height=11}
eop1 = 
mtp_data %>%
  ggplot(aes(x = eop_size, y = eop_size_mm, color = sex)) + 
    geom_point(alpha = 0.8) + 
    labs(
    color = "Gender",
    title = "EOP Size (mm) vs EOP Size Category",
    x = "Category in Chart",
    y = "Continuous EOP size (mm)") +
  theme(axis.text.x = element_text(angle = 45))
eop2 = 
mtp_data %>%
  ggplot(aes(x = eop_visibility_classification, y = eop_size_mm, color = sex)) + 
    geom_point(alpha = 0.8) + 
    labs(
    color = "Gender",
    title = "EOP Size (mm) vs Visibility Classification",
    x = "Category in Chart",
    y = "Continuous EOP size (mm)") +
  theme(axis.text.x = element_text(angle = 45))

fhp = 
mtp_data %>%
  ggplot(aes(x = fhp_category, y = fhp_size_mm, color = sex)) + 
    geom_point(alpha = 0.8) + 
    labs(
    color = "Gender",
    title = "FHP Size (mm) vs FHP Category",
    x = "FHP Category",
    y = "Continuous FHP size (mm)") +
  theme(axis.text.x = element_text(angle = 45))

age = 
mtp_data %>%
  ggplot(aes(x = age_group, y = age, color = sex)) + 
    geom_point(alpha = 0.8) + 
    labs(
    color = "Gender",
    title = "Age vs Age Group Category",
    x = "Age Group Category",
    y = "Age") +
  theme(axis.text.x = element_text(angle = 45))
  
(age + fhp) / (eop1 + eop2)
```


```{r}
age_based = 
  mtp_data %>% 
  filter(sex == "female") %>% 
  count(age) %>% 
  group_by(age > "17" & age < "31", age > "30" & age < "41") %>% 
  summarise(sum = sum(n))

mtp_data %>% 
  filter(sex == "male") %>% 
  count(age)

mtp_data %>% 
  count(age)
```

# Problem 2
```{r}
fhp_p = 
ggplot(data = mtp_data, aes(x = age, y = fhp_size_mm, color = sex)) + 
  geom_point(aes(group = sex), alpha = .5) + 
  geom_smooth() +
  labs(
    color = "Gender",
    title = "FHP size (mm) vs Age",
    x = "Age",
    y = "FHP size (mm)")

gender_total = 
  mtp_data %>% 
  group_by(sex, age_group) %>% 
  summarise(Total = n()) %>% 
  filter(!(age_group == "1"))
  

enlarge = 
  mtp_data %>% 
  filter(eop_size_mm >= 10) %>% 
  group_by(sex, age_group) %>% 
  summarise(enlarge = n()) 

gender_enlarge = 
  left_join(gender_total, enlarge, by = c("age_group","sex")) %>% 
  mutate(enlarge_rate = (enlarge / Total))

enlarge_rate_p = 
ggplot(data = gender_enlarge, aes(x = age_group, y = enlarge_rate, color = sex)) + 
  geom_line(aes(group = sex), alpha = .5) + 
  labs(
    color = "Gender",
    title = "rate of enlarged EOP vs Age Groups",
    x = "Age Groups",
    y = "Rate of Enlarged EOP")

(fhp_p + enlarge_rate_p)
```

```{r}
female = 
  mtp_data %>% 
  filter(sex == "female") %>% 
  filter(!(age_group == "1"))
female_plot = 
  ggplot(data = female, aes(x = eop_size_mm, y = fhp_size_mm, color = age_group)) + 
  geom_point(alpha = .5) +
  geom_smooth() +
  facet_grid(. ~ age_group)

male = 
  mtp_data %>% 
  filter(sex == "male") %>% 
  filter(!(age_group == "1"))
male_plot = 
  ggplot(data = male, aes(x = eop_size_mm, y = fhp_size_mm, color = age_group)) + 
  geom_point(alpha = .5) +
  geom_smooth() +
  facet_grid(. ~ age_group)

female_plot / male_plot
```


