---
title: "p8105_mtp_sl4662"
author: "Siying Li" 
date: "10/25/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(patchwork)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "100%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Introduction 
From previous studies, it is known that development of prominent exostosis young adults’ skulls emanating from external occipital protuberance (EOP). This report is to study wheather the degree of forward head protraction (FHP), sex and age can predict the presence of enlarged EOP (EEOP).

## Results

### Loading and tidying data

First step is to put categorical and continuous variables in appropriate types and fill in missing values. 

```{r}
mtp_data = 
  read_excel("./data/p8105_mtp_data.xlsx", 
             skip = 8) %>% 
  janitor::clean_names() %>% 
  mutate(
    eop_size_mm = replace(eop_size_mm, is.na(eop_size_mm), "0"), 
    eop_size_mm = as.double(eop_size_mm), 
    sex = factor(sex, levels = c("1", "0")), 
    sex = fct_recode(sex, female = "0", male = "1"), 
    age_group = factor(age_group, levels = c("1", "2", "3", "4", "5", "6", "7", "8")),  
    age_group = fct_recode(age_group, 
                           "18-30" = "2", 
                           "31-40" = "3", 
                           "41-50" = "4", 
                           "51-60" = "5", 
                           "60+" = "6", 
                           "60+" = "7", 
                           "60+" = "8"),
    eop_size = factor(eop_size, levels = c("0", "1", "2", "3", "4", "5")), 
    eop_size = fct_recode(eop_size, 
                          "0-5" = "0", 
                          "5-10" = "1", 
                          "10-15" = "2", 
                          "15-20" = "3", 
                          "20-25" = "4", 
                          "25+" = "5"), 
    eop_visibility_classification = factor(eop_visibility_classification, levels = c("0", "1", "2")), 
    eop_visibility_classification = fct_recode(eop_visibility_classification, 
                                               "eop size 0 mm" = "0",
                                               "0 < eop size mm <=5" = "1", 
                                               "eop size >= 5 mm" = "2"),
    fhp_category = factor(fhp_category, levels = c("0", "1", "2", "3", "4", "5", "6", "7")),
    fhp_category = fct_recode(fhp_category, 
                              "0-10" = "0", 
                              "10-20" = "1", 
                              "20-30" = "2", 
                              "30-40" = "3", 
                              "40-50" = "4", 
                              "50-60" = "5", 
                              "60-70" = "6", 
                              "70-80" = "7")
    )

```


Key variables include `age`, `sex`, `eop_size_mm`, `fhp_size_mm`. 

There are `r ncol(mtp_data)` variables and `r nrow(mtp_data)` observations in this dataset. 

After tidying, there are `r nrow(mtp_data)` participants. 

### Age and Gender Distribution

```{r}
age_female = 
  mtp_data %>% 
  filter(sex == "female") %>% 
  group_by(age_group) %>% 
  summarize(female_number = n())

age_male = 
  mtp_data %>% 
  filter(sex == "male") %>% 
  group_by(age_group) %>% 
  summarize(male_number = n())

age_table = 
  left_join(age_female, age_male, by = "age_group") 

knitr::kable(age_table)
```

Number of participants in each group is evenly distrubutted. 


### Issues in Avilable Data

```{r, fig.height=11}
eop1 = 
mtp_data %>%
  ggplot(aes(x = eop_size, y = eop_size_mm, color = sex)) + 
    geom_point(alpha = 0.5) + 
    labs(
    color = "Gender",
    title = "Fig.3 EOP Size (mm) vs EOP Size Category",
    x = "Category in Chart",
    y = "Continuous EOP size (mm)") +
  theme(axis.text.x = element_text(angle = 45))
eop2 = 
mtp_data %>%
  ggplot(aes(x = eop_visibility_classification, y = eop_size_mm, color = sex)) + 
    geom_point(alpha = 0.5) + 
    labs(
    color = "Gender",
    title = "Fig.4 EOP Size (mm) vs Visibility Classification",
    x = "Category in Chart",
    y = "Continuous EOP size (mm)") +
  theme(axis.text.x = element_text(angle = 45))

fhp = 
mtp_data %>%
  ggplot(aes(x = fhp_category, y = fhp_size_mm, color = sex)) + 
    geom_point(alpha = 0.5) + 
    scale_y_continuous(
      breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90)) +
    labs(
    color = "Gender",
    title = "Fig.2 FHP Size (mm) vs FHP Category",
    x = "FHP Category",
    y = "Continuous FHP size (mm)") +
  theme(axis.text.x = element_text(angle = 45)) 

age = 
mtp_data %>%
  ggplot(aes(x = age_group, y = age, color = sex)) + 
    geom_point(alpha = 0.5) + 
    labs(
    color = "Gender",
    title = "Fig.1 Age vs Age Group Category",
    x = "Age Group Category",
    y = "Continuous Age") +
  theme(axis.text.x = element_text(angle = 45))
  
(age + fhp) / (eop1 + eop2)
```

In Fig.1, there are two observations that're categorized in the wrong age group. Such small error in age group would not make an impact on the overall statistics. 

In Fig. 2 and Fig. 3, multiple observations, mostly males, in groups EOP size (20-25mm) and FHP size (40-50mm) are misplaced. 

In Fig. 4, most observations in these two groups are misplaced and making conclusion based on in EOP size 0 mm and EOP size 0-5 mm group will give wrong results. 

## Replot Fig. 3, 4 From Original Paper

```{r}
fhp_p = 
ggplot(data = mtp_data, aes(x = age, y = fhp_size_mm, color = sex)) + 
  geom_point(aes(group = sex), alpha = .5) + 
  geom_smooth() +
  labs(
    color = "Gender",
    title = "Fig. 5 FHP size (mm) vs Age",
    x = "Age",
    y = "FHP size (mm)")

gender_total = 
  mtp_data %>% 
  group_by(sex, age_group) %>% 
  summarise(Total = n()) %>% 
  filter(!(age_group == "1"))
  

enlarge = 
  mtp_data %>% 
  filter(eop_size_mm >= 10) %>% 
  group_by(sex, age_group) %>% 
  summarise(enlarge = n()) 

gender_enlarge = 
  left_join(gender_total, enlarge, by = c("age_group","sex")) %>% 
  mutate(enlarge_rate = (enlarge / Total))

enlarge_rate_p = 
ggplot(data = gender_enlarge, aes(x = age_group, y = enlarge_rate, color = sex)) + 
  geom_line(aes(group = sex), alpha = .5) + 
  labs(
    color = "Gender",
    title = "Fig. 6 Rate of enlarged EOP vs Age Groups",
    x = "Age Groups",
    y = "Rate of Enlarged EOP")

(fhp_p + enlarge_rate_p)
```

From Fig. 5, we can tell that average FHP size of male is always larger than female's. As age goes up, FHP size grows exponentially. In addition, the difference between average and standard deviation both increase as age increases.

From Fig. 6, we know that rate of enlarged EOP among male is always around 30% higher than female. Generally, rate is highest among 18-30 age group. 


## Effect of FHP Size, Age, and Sex on EOP size
```{r}
female = 
  mtp_data %>% 
  filter(sex == "female") %>% 
  filter(!(age_group == "1"))
female_plot = 
  ggplot(data = female, aes(x = eop_size_mm, y = fhp_size_mm, color = age_group)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE, method = lm) +
  facet_grid(. ~ age_group) + 
  theme(legend.position = "none") +
  labs(
    color = "Age Group",
    title = "Fig. 7 Female FHP vs EOP size (mm)",
    x = "EOP Size (mm)",
    y = "FHP Size (mm)") +
  scale_y_continuous(
      breaks = c(0, 25, 50, 75))


male = 
  mtp_data %>% 
  filter(sex == "male") %>% 
  filter(!(age_group == "1"))
male_plot = 
  ggplot(data = male, aes(x = eop_size_mm, y = fhp_size_mm, color = age_group)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE, method = lm) +
  facet_grid(. ~ age_group) + 
  labs(
    color = "Age Group",
    title = "Male FHP vs EOP size (mm)",
    x = "EOP Size (mm)",
    y = "FHP Size (mm)") +
  scale_y_continuous(
      breaks = c(0, 25, 50, 75))

female_plot / male_plot
```

There is positive correlation among all age and sex groups. 

## Comparing results 

### Sample Sizes in Each Age Group

```{r}
age_table = 
  age_table %>% 
  mutate(total = female_number + male_number) %>% 
  select(age_group, total)

knitr::kable(age_table)
```

From the data set given, numbers of participants in each age group doesn't have consistent numbers with the author’s stated sample size. 

### FHP Size Data

```{r}
mtp_data %>% 
  drop_na(fhp_size_mm) %>% 
  group_by(sex) %>% 
  summarise(mean = mean(fhp_size_mm), sd = sd(fhp_size_mm)) %>% 
  knitr::kable()
```

When not replacing the missing value with 0, above data is consistent with author's calculation if we round the numbers. 

### Prevalence of EEOP

```{r}
eeop = 
  mtp_data %>%
  filter(eop_size_mm > 10) %>% 
  nrow()

eeop / (nrow(mtp_data)) * 100
```

EEOP is the ones with EOP exceeded 10mm in size. I used `eop_size_mm` to calculate prevalence of EEOP to be `r eeop / (nrow(mtp_data) - 1) * 100`% of the study population. They are similar (not consistent).

### FHP Among 60+ Age Group

```{r}
fhp_percent = 
mtp_data %>%
  filter(age > 60 & fhp_size_mm > 40) %>% 
  nrow()

over_six = 
  mtp_data %>%
  filter(age > 60) %>% 
  nrow()

(fhp_percent / over_six) * 100
```

From the data set given, FHP >40 mm observed frequently in the over 60s cases is `r (fhp_percent / over_six) * 100`%, while author got 34.5%. They are similar (not consistent). Also, according to Fig. 5, FHP size goes up as age increases. 

# Discussion

Based on the given data set, some of the results are inconsistent with author’s results due to misplacement of continuous observations into wrong groups. Solely based on the data, I cannot conclude that there is causal relationship between cell phone and “horn” growth. I think it would be more helpful to look at EPO size from of people use cell phone and people who do not. 