---
title: "p8105_mtp_sl4662"
output: github_document
---

Some age group does not match the age. Age group 1 is not suppose to appear, and there is an age 45, with age group 1. 

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(readxl)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
mtp_data = 
  read_excel("p8105_mtp_data.xlsx", 
             skip = 8) %>% 
  janitor::clean_names() %>% 
  mutate(eop_size_mm = replace(eop_size_mm, is.na(eop_size_mm), "0")) %>% 
  mutate(age_group = factor(age_group, levels = c("1", "2", "3", "4", "5", "6", "7", "8"))) %>%
  mutate(sex = factor(sex, levels = c("1", "0"))) %>% 
  mutate(eop_visibility_classification = factor(eop_visibility_classification, levels = c("0", "1", "2"))) %>% 
  mutate(eop_size= factor(eop_size, levels = c("0", "1", "2", "3", "4", "5"))) %>% 
  mutate(fhp_category= factor(fhp_category, levels = c("0", "1", "2", "3", "4", "5", "6", "7")))
# levels(mtp_data$age_group) = c("1", "20+", "30+", "40+", "50+", "60+", "60+", "60+")

age_female = 
  mtp_data %>% 
  filter(sex == "0") %>% 
  group_by(age_group) %>% 
  summarize(female_number = n())

age_male = 
  mtp_data %>% 
  filter(sex == "1") %>% 
  group_by(age_group) %>% 
  summarize(male_number = n())

left_join(age_female, age_male, by = "age_group") %>% 
  mutate(age = c("10+", "20+", "30+", "40+", "50+", "60+", "70+", "80+")) %>% 
  select(-age_group) %>% 
  select(age, everything()) %>% 
  knitr::kable()
```

This dataset is consist of many different variables, and some variables have conclusive variables following them to classify the results into different groups. 

Key variables include :

There are `r ncol(mtp_data)` variables and `r nrow(mtp_data)` observations in this data set. There is a total of `r nrow(mtp_data)` participants. 


